Efficient spherical Convolutional Neural Networks with Healpix sampling for cosmological applications

### 2018-03-07 scnn.tex
update structure: method section, separate related work

\subsection{Motivation}
\assign{Tomek}

@@

\subsection{Potential applications}
	\assign{Tomek, Nathanael, Michael}

@@

Finally, not that those neural networks are not restricted to the sphere and can be applied to any problem where we have data on a graph, such as social, biological or infrastructure networks [some citations, e.g. brain Alzeihmer, particle physics].
% the convolution is not restricted to the sphere, the coarsening/pooling is

\subsection{Related work}
\assign{Nathanael, Michael}
\label{sec:related}
\cite{cohen2017convolutional} 
\cite{cohen2018spherical}
\cite{boomsma2017spherical}
\cite{defferrard2016convolutional}

\nati{+ add other graph CNN approaches \cite{...}}

@@

@@

Related to this, convolutional neural networks have been defined on manifolds and have achieved impressive results on shapes [Bronstein]. They however too depend on an orientation, which a sphere do not possess.

\section{Spherical convolution on sphere using a graph}
\assign{Nathanael}

@@

\paragraph{Graph creation.}
As a hard constraint resulting from the cosmological applications, we are bound to use the HealPix sampling~\cite{gorski2005healpix} and we create a weighted undirected graph where each pixel is a node (vertex) that is connected to his $8$ or $7$ closest neighboors\footnote{For some pixels, the $8^{th}$ nearest neighbor is not well defined.}. Given the set of nearest neighbors, we define the weight matrix $W$ using the following scheme
\begin{equation}
W[i,j]=\begin{cases}
e^{-\frac{\|x_i-x_j\|_2^2}{\sigma^2}} & \text{if pixels $i$ and $j$ are connected, and}\\
0 & \text{otherwise.}\\
\end{cases}
\end{equation}
Here $x_i$ is a 3-dimensional vector encoding the coordinate of the pixels $i$ on the sphere and $\sigma$ is the mean of $\|x_i-x_j\|_2$ over all connected pixels $i$ and $j$. The degree of a node (or a pixel) is defined as $d_i = \sum_j W[i,j]$. The degree matrix $D$ is the diagonal matrix where $D_ii=d_i$.

\paragraph{Graph Fourier Transform and spherical harmonics.}
The graph normalized graph Laplacian defined as $L = I - D^{-1/2} W D^{-1/2}$ is a second order differential operator that can be used to define the graph Fourier basis.

@@

\section{Efficient spherical CNN}
\assign{Michael}

@@

\section{Experiments}
\assign{Nathanael, Tomek}

@@

>>>>

%\subsection{Motivation}
\assign{Tomek}

@@

\subsection{Potential applications}
	\assign{Tomek, Nathanaël, Michaël}

@@

Finally, not that those neural networks are not restricted to the sphere and can be applied to any problem where we have data on a graph, such as social, biological or infrastructure networks [some citations, e.g. brain Alzeihmer, particle physics, computer graphics].
% the convolution is not restricted to the sphere, the coarsening/pooling is

\section{Related work}
\label{sec:related}

\subsection{Spherical convolutions}
\assign{Nathanaël}

\cite{cohen2017convolutional} 
\cite{cohen2018spherical}
\cite{boomsma2017spherical}

\subsection{Convolutions on graphs}
\assign{Michaël}

\todo{other approaches? GNNs, Kipf first order approx, message passing}

@@

In this work, we are using the graph CNN formulation introduced in \cite{defferrard2016convolutional}.

@@

\subsection{Convolutions on manifolds}

Related to this, convolutional neural networks have been defined on manifolds and have achieved impressive results on shapes [Bronstein]. They however too depend on an orientation, which spheres do not possess.

\subsection{Convolutions on point clouds?}

PointNet and co. Related but we are loosing the structure. Also coarsening.

\section{Method}

The gist of our method is to define the convolution on a sphere using a graph. The graph is here seen as a discrete approximation of the sphere $S^2$, a 2-dimensional manifold embedded in $\mathbb{R}^3$.

@@

\subsection{Graph creation}
\assign{Nathanaël}

As a hard constraint resulting from the cosmological applications, we are bound to use the HealPix sampling~\cite{gorski2005healpix} and we create a weighted undirected graph where each pixel is a node (vertex) that is connected to his $8$ or $7$ closest neighboors\footnote{For some pixels, the $8^{th}$ nearest neighbor is not well defined.}. Given the set of nearest neighbors, we define the weight matrix $W$ using the following scheme
\begin{equation}
W[i,j]=\begin{cases}
e^{-\frac{\|x_i-x_j\|_2^2}{\sigma^2}} & \text{if pixels $i$ and $j$ are connected, and}\\
0 & \text{otherwise.}\\
\end{cases}
\end{equation}
Here $x_i$ is a 3-dimensional vector encoding the coordinate of the pixels $i$ on the sphere and $\sigma$ is the mean of $\|x_i-x_j\|_2$ over all connected pixels $i$ and $j$. The degree of a node (or a pixel) is defined as $d_i = \sum_j W[i,j]$. The degree matrix $D$ is the diagonal matrix where $D_ii=d_i$.

\subsection{Coarsening}
\assign{Nathanaël}

How we get coarser or finer sampling of the sphere with Healpix. The $2^k \cdot 12$ points.

\subsection{Graph Fourier basis and spherical harmonics}
\assign{Nathanaël}
% TODO: basis or transform?

The graph normalized graph Laplacian defined as $L = I - D^{-1/2} W D^{-1/2}$ is a second order differential operator that can be used to define the graph Fourier basis.

@@

\subsection{Convolutions on graphs}
\assign{Nathanaël, Michaël}

With the Fourier basis. Because we don't have an FFT, that costs $O(n^3)$ for the eigendecomposition, plus $O(n^2)$ for the transform (to be done for each forward and backward pass).

\subsection{Efficient convolutions}
\assign{Michaël}

@@

\section{Experiments}
\assign{Nathanaël, Tomek}

@@

\section{Conclusion}
\assign{Nathanaël, Tomek, Michaël}
